---
title: "UN Speeches Wordcloud"
output: html_document
date: "2026-08-12"
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE}
library(tidyverse)
library(readtext)
library(tidytext)
library(lubridate)
library(ggwordcloud)
```


```{r}
UN_speeches <- tibble(readtext("dataverse_files/UNGDC_1946-2025/TXT/", 
         docvarsfrom = "filenames", 
         docvarnames = c("country", "sessions", "year"), 
         dvsep = "_")) 
```



```{r}
UN_speeches |> 
  count(country, sort = TRUE)
```

# Filter and clean data

```{r}
UN_speeches |> 
  #Select countries
  filter(country %in% c("USA","GBR", "RUS", "DEU", "POL", "FRA", "CHN", "BRA", "YUG")) |>
    #select years
  filter(year >= 1989, year <= 2000) |> 
  #Remove digits
  mutate(text = str_remove_all(text, "\\d+")) -> UN_speeches_clean
```

```{r}
UN_speeches_clean |> 
  unnest_tokens(word, text) -> UN_speeches_clean_tidy
```



```{r}
UN_speeches_clean_tidy |> 
  count(country, word) |> 
  bind_tf_idf(word, country, n) |> 
  arrange(desc(tf_idf)) -> UN_speeches_clean_tfidf
```



```{r}
UN_speeches_clean_tfidf
```
# Stopwords
```{r}
country_stopwords <- tibble(word = c("australia's", "australia’s", "australians", "uk", "america", "hong", "kong", "britain", "britain’s", "germany", "canada", "canadian", "canada's", "canadians", "France", "France’s", "polish", "americans", "germany's", "german", "germans", "british", "kong's", "Canada´s", "english", "bonn", "â", "france", "france’s", "twelve", "canada’s", "Canada’s", "franc", "france's", "russian", "pearson", "americas", "walk", "congress", "cfe", "b", "poland", "brazilian", "taiwan", "china", "china's", "brazil", "russia", "China", "albania", "Brazilians", "Brazil's", "chinese", "China's", "brazilians", "brazil's", "chile", "ninth", "cambodian", "china’s", "brazil’s", "poland’s", "french", "cardoso", "albanian"))
```
# Figure 1
```{r}
#Code for figure 1
UN_speeches_clean_tfidf |> 
#we uses the clean_tfidf object and pipes anti_join, witch sorts out the stopwords
  anti_join(country_stopwords) |> 
#arrange the values decending
  arrange(desc(tf_idf)) |> 
#modifying the word-column with mutate() to the plot function, with the most common words first by the rev() function.
  mutate(word = factor(word, levels = rev(unique(word)))) |> 
#group by country
  group_by(country) |> 
#selecting the eight most common words after the filtering for each country
  top_n(8) |>  
#plotting and labeling
  ggplot(aes(label = word, size = tf_idf, color = tf_idf)) +
#plotting a wordcloud
  geom_text_wordcloud_area() +
#scaling the size of the words
  scale_size_area(max_size = 13) +
  theme_minimal() +
#ordering the look of the nine wordclouds
  facet_wrap(~country, ncol = 3, scales = "free") +
  scale_color_gradient(low = "darkblue", high = "blue") +
#giving names so the figure is easy to understand.
  labs(
      title = "UN Speeches: most important words pr. country in the timespan from year 1989-2000",
       subtitle = "Importance determined by term frequency (tf) - inversed document frequency(idf)",
      caption = "Data from United Nations General Debate Corpus")
```

# Figure 2
```{r}
UN_speeches_clean_tfidf |>  
  anti_join(country_stopwords) |> 
  group_by(country) |>  
#Here we only have room for the seven most common words for the bar graphs to be readable.
  slice_max(tf_idf, n = 7) |>  
  ungroup() |>  
  ggplot(aes(
    x = reorder(word, n),
    y = n
  )) +
  geom_col() +
  coord_flip() +
  facet_wrap(~country, ncol = 3, scales = "free") +
  theme_minimal() +
  labs(
    title = "Country-characteristic words in UN speeches in the period 1989-2000",
    x = "Words",
    y = "Number of times written in a speech by country"
  )
```

# Defining the eight topwords and viewing them
```{r}
Top_words <- UN_speeches_clean_tfidf |> 
  anti_join(country_stopwords) |> 
  arrange(desc(tf_idf)) |> 
  mutate(word = factor(word, levels = rev(unique(word)))) |> 
  group_by(country) |> 
  top_n(8)
Top_words
```

# Finding specific words with the kwic-function:
```{r, message=FALSE}
library(quanteda)
```
```{r}
UN_speeches_quanteda <- UN_speeches |> 
  filter(year %in% C(1989: 2000))

corp <- corpus(UN_speeches_quanteda, docid_field = "doc_id", text_field= "text")
```
```{r}
#skal der være komma og punktum i dataen. Ja, vi vil gerne se i konteksten hvornår en ny sætning starter. Derfor FALSE nedenfor
specific_term_data <- kwic(tokens(corp, remove_punct = FALSE),
#selecting the keyword to 'multipolar'
     pattern = phrase("multipolar"),
#viewing the seven word before and after the keyword
     window = 7)
```
# Defining the specific term multipolar
We build in coloums by column with the relevant data. The country and year it is sayd, the seven words before, the keyword and the seven words after. 
```{r}
specific_term <- 
  data.frame(
  country_year = specific_term_data$docname,
  pre = specific_term_data$pre,
  keyword = specific_term_data$keyword,
  post = specific_term_data$post)
specific_term
```

```{r}
#defining the specific term so that next time we uses the function, our data for the keyword in this case - 'multipolar' will be stored in the object multipolar and not forgotten.
multipolar <- specific_term
```

#The same goes for 'end of history', 'liberal democracy', 'historical Proess' and 'unipolar'. First unipolar:
```{r}
specific_term_data <- kwic(
  tokens(corp, remove_punct = FALSE),
  pattern = phrase("unipolar"),
  window = 7)

specific_term <- 
  data.frame(
  country_year = specific_term_data$docname,
  pre = specific_term_data$pre,
  keyword = specific_term_data$keyword,
  post = specific_term_data$post)
specific_term

```
#End of history
```{r}
specific_term_data <- kwic(
  tokens(corp, remove_punct = FALSE),
  pattern = phrase("end of history"),
  window = 7)

specific_term <- 
  data.frame(
  country_year = specific_term_data$docname,
  pre = specific_term_data$pre,
  keyword = specific_term_data$keyword,
  post = specific_term_data$post)
specific_term
```
#Liberal democracy and figure:
```{r}
liberal_democracy <- kwic(tokens(corp, remove_punct = TRUE),
     pattern = phrase("liberal democracy"),
     window = 5)
country_counts <- liberal_democracy |>
  as.data.frame() |> mutate(country = str_extract(docname, "^[A-Z]{3}")) |>
  count(country, name = "phrase_count")

ggplot(country_counts, aes(
  x = reorder(country, phrase_count),
  y = phrase_count
)) +
  geom_col(fill = "blue") +
  coord_flip() +
  labs(
    x = "Country",
    y = 'Number of times "liberal democracy" is said',
    title = 'Mentions of "liberal democracy" by country'
  ) +
  theme_minimal()
```

#historical process and figure:
```{r}
historical_process <- kwic(
  tokens(corp, remove_punct = TRUE),
  pattern = phrase("Historical Process"),
  window = 5)
#Here we have used AI to find the str_extract function, for more information see our GAI-declaration.
country_counts <- historical_process |>
  as.data.frame() |> mutate(country = str_extract(docname, "^[A-Z]{3}")) |>
  count(country, name = "phrase_count")

ggplot(country_counts, aes(
  x = reorder(country, phrase_count),
  y = phrase_count
)) +
  geom_col(fill = "blue") +
  coord_flip() +
  labs(
    x = "Country",
    y = 'Number of times "historical process" is said',
    title = 'Mentions of "historical process" by country'
  ) +
  theme_minimal()
```

